Recurrent event analysis in the presence of real-time high frequency data via random subsampling

### 2020-02-28 write-up/fda-recurrent.tex
Finished proof of asymp normality

\begin{proof}[Proof of Asymptotic Normality]
Consider a Taylor-expansion of $\tilde U_n (\theta)$ centered at $\theta_0$ when $\hat \theta_n \in \Theta_0$.  We can write as
\begin{align*}
\tilde U_n (\hat \theta_n) &= \tilde U_n (\theta_0) -  (\hat \theta_n - \theta_0) \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} - (\hat \theta_n - \theta_0) \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \\
0 &= \tilde U_n (\theta_0) + (\hat \theta_n - \theta_0) \left[ \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \right] \\
n^{1/2} (\hat \theta_n - \theta_0) &= \left[ \frac{1}{n} \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \frac{1}{n} \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \right]^{-1} n^{-1/2} \tilde U_n (\theta_0)
\end{align*}
where~$\theta^\star$ is on the line segment between $\hat \theta_n$ and $\theta_0$. First, the term
$$
\frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star}
\overset{P}{\to} \Xi(\theta_0)
$$
under conditions VI.1.1(A)--(E) in~\cite{Andersen1993} and Assumptions~\ref{E1}--\ref{E4}.  To see this, note the first term converges to $\Xi(\theta_0)$ while the second term is bounded in probability by $p M \| \theta^\star - \theta_0 \|$ for some finite constant~$M$ independent of~$\theta^\star$ where $\| \cdot \|$ is the Euclidean norm.  This goes to zero since $\| \theta^\star - \theta_0 \| \to 0$.

Now $\tilde U_n (\theta_0 )$ can be decomposed into three terms as in~\eqref{eq:decomp}. Here, we assume $K_x \to \infty$ so the final term is negligible.  This leaves two terms
$$
U_n (\theta_0) = \sum_{i=1}^n
$$ and
$$
n^{-1} (\hat U_n (\theta_0) - U_n (\theta_0) ) = \frac{1}{n} \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(s; \theta_0)}{\pi_i (s) + h(s; \theta_0)} dM_i (s)
$$
\end{proof}

\begin{proof}[Proof of Proposition~\ref{prop:optimal}]
Consider the weighted estimating equations given by~\eqref{eq:approxscore}.
Then
\end{proof}

>>>>

\begin{proof}[Proof of Asymptotic Normality]
Consider a Taylor-expansion of $\tilde U_n (\theta)$ centered at $\theta_0$ when $\hat \theta_n \in \Theta_0$.  We can write as
\begin{align*}
\tilde U_n (\hat \theta_n) &= \tilde U_n (\theta_0) -  (\hat \theta_n - \theta_0) \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} - (\hat \theta_n - \theta_0) \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \\
0 &= \tilde U_n (\theta_0) + (\hat \theta_n - \theta_0) \left[ \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \frac{1}{2} \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \right] \\
n^{1/2} (\hat \theta_n - \theta_0) &= \left[ \frac{1}{n} \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \frac{1}{2n} \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star} \right]^{-1} n^{-1/2} \tilde U_n (\theta_0)
\end{align*}
where~$\theta^\star$ is on the line segment between $\hat \theta_n$ and $\theta_0$. First, the term
$$
\frac{1}{n} \frac{\partial}{\partial \theta} \tilde U_n (\theta) \mid_{\theta=\theta_0} + \frac{1}{2n} \sum_{l=1}^p (\hat \theta_{ln} - \theta_{l0}) \frac{\partial^2}{\partial \theta \partial \theta_l} \tilde U_n (\theta) \mid_{\theta=\theta^\star}
\overset{P}{\to} \Xi(\theta_0)
$$
under conditions VI.1.1(A)--(E) in~\cite{Andersen1993} and Assumptions~\ref{E1}--\ref{E4}.  To see this, note the first term converges to $\Xi(\theta_0)$ while the second term is bounded in probability by $p M \| \theta^\star - \theta_0 \|$ for some finite constant~$M$ independent of~$\theta^\star$ where $\| \cdot \|$ is the Euclidean norm.  The second term therefore goes to zero since $\| \theta^\star - \theta_0 \| \to 0$ as $n \to \infty$.

We now consider $n^{-1/2} \tilde U_n (\theta_0 )$. Recall $\tilde U_n (\theta_0)$ can be decomposed into three terms as in~\eqref{eq:decomp}. Here, we assume $K_x \to \infty$ so the final term is negligible.  This leaves two terms
$$
U_n (\theta_0) = \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(u; \theta)}{h (u;\theta)} dM_i^{(e)} (t)
$$
where~$N_i^{(e)} (t)$ and $N_i^{(s)} (t)$ are the orthogonal event and subsampling counting processes with jumps at times $t \in \bfT_i$ and $t \in \bfD_i$ respectively.
and
\begin{align*}
(\hat U_n (\theta_0) - U_n (\theta_0) )
&= \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(s; \theta_0)}{\pi_i (s) + h(s; \theta_0)} dM_i (s) \\
&= \sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(s; \theta_0)}{\pi_i (s) + h(s; \theta_0)} dM^{(e)}_i (s) +
\sum_{i=1}^n \int_0^\tau \frac{h^{(1)}(s; \theta_0)}{\pi_i (s) + h(s; \theta_0)} dM_i^{(s)} (s) \\
&= \sum_{i=1}^n \int_0^\tau (1-w(s;\theta_0)) \cdot \frac{h^{(1)}(s; \theta_0)}{h_i (s;\theta_0)} dM_i^{(e)} (s) +
\sum_{i=1}^n \int_0^\tau w(s;\theta_0) \cdot \frac{h^{(1)}(s; \theta_0)}{\pi_i (s)} dM_i^{(s)} (s)
\end{align*}
Taking difference with $U_n(\theta_0)$, the first term becomes
\[
\sum_{i=1}^n \int_0^\tau w(s;\theta_0) \cdot \frac{h^{(1)}(s; \theta_0)}{h_i (s;\theta_0)} dM_i^{(e)} (s)
\]

These are two orthogonal square integrable martingales with quadratic variation given by
\[
\sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \left[ \frac{h^{(1)}(s; \theta_0)}{h_i (s;\theta_0)} \right] \times \left[ \frac{h^{(1)} (s;\theta_0)}{h_i (s;\theta_0)} \right]^\top h_i (s) ds
\]
and
\[
\sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \left[ \frac{h^{(1)}(s; \theta_0)}{\pi_i (s)} \right] \times \left[ \frac{h^{(1)} (s;\theta_0)}{\pi_i (s)} \right]^\top \pi_i (s) ds
\]
respectively.  Using the definition $w(s;\theta_0) = \pi_i(s) / (\pi_i(s)+h_i(s;\theta_0))$, we have
\begin{align*}
&\frac{1}{n} \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \left[ \frac{h^{(1)}(s; \theta_0)}{h_i (s;\theta_0)} \right] \times \left[ \frac{h^{(1)} (s;\theta_0)}{h_i (s;\theta_0)} \right]^\top h_i (s) ds \\
&+
\frac{1}{n} \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \left[ \frac{h^{(1)}(s; \theta_0)}{\pi_i (s)} \right] \times \left[ \frac{h^{(1)} (s;\theta_0)}{\pi_i (s)} \right]^\top \pi_i (s) ds \\
=& \frac{1}{n} \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h_i (s;\theta_0)} \left[ 1 + \frac{h(s;\theta_0)}{\pi_i (s)} \right] ds \\
=& \frac{1}{n} \sum_{i=1}^n \int_0^\tau w^2(s;\theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h_i (s;\theta_0)} \left[ \frac{\pi_i (s) + h(s;\theta_0)}{\pi_i (s)} \right] ds \\
=& \frac{1}{n} \sum_{i=1}^n \int_0^\tau w(s;\theta_0) \frac{h^{(1)}(s; \theta_0) [h^{(1)} (s;\theta_0)]^\top}{h_i (s;\theta_0)} ds
\to \Xi (\theta_0).
\end{align*}
Returning to the Taylor-expansion, under conditions VI.1.1(A)--(E) in~\cite{Andersen1993} and Assumptions~\ref{E1}--\ref{E4}, we can apply Rebolledo's martingale central limit theorem so that
$$
\sqrt{n} (\hat \theta_n - \theta_0) \overset{D}{\to} N(0, \Xi(\theta_0) ^{-1} \times \Xi(\theta_0) \times \Xi(\theta_0) ^{-1} )
$$
which completes the proof.
\end{proof}

\begin{proof}[Proof of Proposition~\ref{prop:optimal}]
The above proof establishes asymptotic normality for all
The main additional assumption is

The argument

\end{proof}
